Manuscript

### 2025-04-09 Manuscript_MD_ScoreReLiability.qmd
init manuscript V6 update

Large-scale collaborative replication efforts have sparked discussions
surrounding the replicability of psychological phenomena. For attempts
to estimate a phenomenon’s replicability or to predict whether a future
replication will replicate successfully, heterogeneity is a crucial
parameter that needs to be assessed. In the meta-analytic context,
heterogeneity describes the variation in effect sizes, free of sampling
error, if these are not constant across replications. Therefore, the
presence of heterogeneity implies that some replications of a single
phenomenon may be successful, while others are not. If heterogeneity
grows, meaning the phenomenon’s effect size varies more strongly for
unexplained reasons, the probability of observing an effect size around
zero or even in the negative space grows larger as well. If we know the
mean size and heterogeneity of a phenomenon’s effect, we would
theoretically be able to establish an expected replication rate
(reference).

Similarly, it has been argued that heterogeneity in effect sizes is an
indicator of the theory’s “completeness” surrounding the phenomenon.
Linden and Hönekopp argue that “low (as opposed to high) heterogeneity
reflects a more advanced understanding of the subject matter being
studied” (2021, p.2). Alternatively, Schuetze and von Hippel (2023)
argue that heterogeneity in effects is an indicator of a vague, poorly
specified theory.

Large-scale attempts of direct replications, using identical protocols,
such as the Many Labs studies or Registered Replication Reports
(references), for the first time allow researchers to estimate
heterogeneity undistorted by selection processes induced by publication
bias. In re-analyses of these studies, Olsson-Collentine et al. (2020)
identify a positive correlation between a phenomenon’s effect size and
its heterogeneity. Similarly, van Erp et al. (2017) or Stanley et al.
(2018) estimate strong degrees of heterogeneity across conceptual
replications in Psychology. In a separate re-analysis of large scale
direct replications, Renkewitz et al. (in preparation) identify
substantial heterogeneity in almost all projects where a non-zero effect
could be identified. This aligns with the correlation found by
Olsson-Collentine et al. (2020).

One source for this heterogeneity, which is not a typical moderator tied
to a specific theory, might be impairments in score reliability which
vary in their extent across replications. This is an argument put
forward repeatedly by original researchers and replicators alike.
Literature discussing the role of score reliability for standardized
effect sizes claims that differences in score reliability should
inflate, and therefore increase, heterogeneity. Over the following
pages, we assess whether score reliability can in fact be made
responsible for heterogeneity identified across phenomena. Even though
only a limited number of phenomena allow for such an analysis, we find
no empirical evidence for the argument that differences in score
reliability inflate effect size heterogeneity. Therefore, in the last
part of this text we turn towards an alternative discussion using
analytical arguments on how differences in measurement quality affect
standardized effect sizes and their differences.

@@

Here, MD refers to the difference between two groups of interest, while
$\sigma_X$ is the total pooled standard deviation. Since $d$ constitutes
an observed effect size it is affected by sampling error. In terms of
heterogeneity, we are not interested in the variance of $d$, but the
variance of the true, underlying ES $\delta$. The standard error of
Cohen’s $d$, which is used to weight individual estimates in a
random-effects meta-analysis (reference) is computed as defined in
@eq-SEd.

@@

Therefore, in the following section we assess to what extent this
intuition is supported by empirical evidence. The re-analysis of
archival data serves to shed light on the following two research
questions central to this text:

-   Can score reliability be used to explain heterogeneity in
    standardized effect sizes?

-   How does score reliability affect ES heterogeneity?

@@

The ManyLabs projects were collaborative efforts to replicate several
psychological phenomena across different research sites, employing
identical protocols. From five published projects, the data for the
first three and the fifth Many Labs projects was made publicly available
when the DRIPHT repository was set up (references). The Registered
Replication Reports are similar collaborative efforts, with the sole
distinction that for each report a single phenomenon was replicated
several times. The Registered Reports 3-10 were added to the DRIPHT
repository, as they employed experimental designs with at least two
groups. Lastly, some projects from the Psychological Science Accelerator
were added to the DRIPHT repository. Different to Many Labs or
Registered Replication Projects, these collaborative efforts, while also
distributed across the globe, do not formally attempt to perform direct
replications, but focus on original research or conceptual replications.
However, as for each project, across all sites the same protocol is used
and the subjects are distributed across different countries, a data
structure similar to that of a multi-site replication study emerges.
What makes all these collaborative efforts valuable for this manuscript
is the fact that all phenomena collected in the DRIPHT-repository are
making use of two-group designs and are therefore easily assessed using
standardised effect sizes.

While replication data on a decent number of psychological phenomena is
available in the DRIPHT-repository, to demonstrate how incorporating
score reliability affects heterogeneity in ES, selected phenomena need
to fulfil three conditions: (1) As we focus on the use of ES $d$, the
phenomenon needs to be assessed using a two-group design. This is the
case for all projects in the DRIPHT repository; (2) Score reliability
estimates can be derived, this implies that the phenomenon needs to be
measured using either several indicators forming a single scale or by
repeatedly measuring across several timepoints; (3) Score reliability
can only attenuate effect sizes that are not zero in the first place.
Therefore, we focus on phenomena where meta-analytic mean ES can be
statistically distinguished from zero.

As all phenomena in the collection fulfil the first condition, 50
phenomena have been catalogued where the effect is studied by comparing
the outcome across two separate groups. From the collection of 50
phenomena, 21 fulfil the second condition, so that estimates of score
reliability can be derived. For the 21 phenomena, the dependent variable
used to construct the effect size is measured by employing more than a
single indicator. At the same time, however, this implies that the
remaining 29 phenomena did not make use of more than a single indicator
to measure the dependent variable. Therefore, for the majority of
phenomena it is not possible to assess in how far measurement quality is
sufficiently high in terms of e.g. internal consistency or low random
error variance.

From the 21 phenomena that employed more than a single indicator, 19
made use of Likert-style items, where respondents would indicate their
agreement to some kind of statement. The scale-length varied from 3 to
10-points. The items measuring the remaining 2 phenomena were coded
dichotomously, as responses were right or wrong. While the majority of
designs make use of some form of control-treatment manipulation,
randomly assigning participants to a condition (20 phenomena), some
phenomena refer to the effect of pre-existing differences, e.g.
biological sex (2 phenomena). Subsequently, those phenomena need to be
identified where the meta-analytic mean effect size is statistically
significant different from zero. This will be done as a preliminary step
in this manuscript’s analysis procedure. Detailed information concerning
the different phenomena and how they were measured can be found at
\[osf-link\], while a brief summary can be found in Table 1.

@@

For each replication of a phenomenon, Cohen’s $d$ is computed as an
estimate of raw ES, according to @eq-d, including its standard error
(@eq-SEd). Subsequently, a random-effects meta-analysis is performed
using metafor version X.XX. To estimate meta-analytic mean ES and
heterogeneity, the REML-estimator is used (ref). To identify which
phenomena pass criterion (3), using a Wald-type significance test we
assess for which phenomenon the meta-analytic mean ES is significantly
different from zero. For this hypothesis test, as all other hypothesis
tests in this manuscript, a one-tailed significance level of .05 is
used.

Additionally, estimates of score reliability are derived. While not
without criticism (references), Cronbach’s Alpha is used to estimate
score reliability. It is often noted that tau-equivalence is an
unrealistic assumption to hold against real-world data, implying that
other estimates of score reliability, such as McDonald’s Omega
(reference), Guttman’s Lambda 2 or 4 (reference), or the Greatest Lower
Bound (reference) might be better suited. However, all scales employed
in this project are scored by computing the simple mean or sum of
responses across items for each individual. Computing the mean or sum of
several items implicitly assumes tau-equivalence, as each item
contributes equally to the individual’s test score. Therefore,
Cronbach’s Alpha is the better choice to estimate score reliability, as
it avoids a mismatch in assumptions between how the score is computed
and how score reliability is estimated. Using these estimates of
Cronbach’s Alpha, the individual estimates of Cohen’s d are corrected,
according to @eq-dc, including the corrected standard errors (@eq-SEdc).
Thereby, estimates of ES are computed, which are corrected for imperfect
reliability. Again, a random-effects meta-analysis is run using metafor,
generating an estimate of heterogeneity in corrected ES. Different
indicators of heterogeneity are readily available, such as $I^2$, $H^2$
or the coefficient of variation. However, the descriptions found in H&S
and W&D discuss the absolute amount of heterogeneity in terms of
variance ($\tau^2$) or standard deviation ($\tau$). These parameters
discuss the variability of the standardized ES in the population in
terms of variance or standard deviation. Therefore, we also make use of
absolute heterogeneity here.

Subsequently, in order to assess whether ES heterogeneity was indeed
reduced by the attenuation correction procedure, the estimates of
heterogeneity in uncorrected ES and corrected ES are compared. Using
Cochran’s Q-test, we assess whether the estimates of heterogeneity are
statistically significantly different from 0. Only for those projects
where we have sufficient confidence that the individual estimates of
heterogeneity are larger than zero can we actually begin to interpret
whether there was any change.

@@

These procedures are followed separately for each phenomenon. The
statistical programming language R, Version X.XXX (reference) is used
for all data manipulation and statistical analysis.

### Results - Which phenomena pass criterion 3)?

[Equations @eq-d], [-@eq-SEd], [-@eq-dc], and [-@eq-SEdc] were used to
generate estimates of standardized ES, both uncorrected and corrected,
with their corresponding standard errors. Generally, this leads to
larger (absolute) ES and larger standard errors. As an example from the
21 selected phenomena, the results from this procedure on Nosek’s
Explicit Art sex differences phenomenon are displayed in the forest plot
in @fig-forest. Here, grey dots represent the uncorrected ES for each
sample, while black dots represent the corrected ES in each sample. The
bars surrounding these dots show the respective 95%-Cis.

```{r, echo = FALSE}
#| label: fig-forest
#| fig-cap: Forest Plot Nosek_Explicit_Art
forest_plot_rel(ES_rma_df[which(ES_rma_df$MASC == "Nosek_Explicit_Art"),], agg_L[[which(MASC_names == "Nosek_Explicit_Art")]])
```

@@

@tbl-meanES demonstrates that what we observed in Figure 1 holds across
all 21 phenomena. All meta-analytic effects are larger in their absolute
size after applying an attenuation correction procedure. Similarly, the
uncertainty quantified in the estimate's standard error is larger.
Additionally, @tbl-meanES highlights which phenomena pass criterion (3)
– the effect size must be statistically significantly distinguishable
from zero. 12 phenomena pass this criterion, highlighted in boldface. It
may be remarked that the attenuation correction procedure, in this data,
appears to have no influence on whether a phenomenon passes the
criterion. The differences in p-value are rather small and lead to no
difference in conclusion, regarding a significance level of .05.

```{r, echo = FALSE}
#| label: tbl-meanES
#| tbl-cap: Meta-analytic average  ES
knitr::kable(x = read.csv(here("Tables/Mean_ES.csv")))
```

### Results - Estimates of heterogeneity

@tbl-tauES contains the results from performing an RG-MA on the
Cronbach's Alpha coefficients of the scales used to measure the 12
remaining phenomena. Additionally, it contains estimates of
heterogeneity for both uncorrected ES ($\delta_0$) and corrected ES
($\delta$), which are central to assess the research questions posed in
section XXX. Most importantly, @tbl-tauES demonstrates that for all but
one of the phenomena, the attenuation correction procedure barely has an
impact on the extent of ES heterogeneity. For all phenomena but
Srull_Behaviour_Hostility, the heterogeneity in ES before and after the
correction changes very little. This implies that for 11 out of 12
phenomena, differences in score reliability can not be made responsible
for heterogeneity in standardized effect sizes. Only for the
Srull_Behaviour_Hostility phenomenon, accounting for score reliability
actually does reduce, and therefore explain, ES heterogeneity.

@@

*Note:* $\hat{\mu}_{\rho_{XX'}}$ is the back-transformed meta-analytic
estimate of mean score reliability; $\tau{\mu}_{\rho_{XX'}}$ is the
back-transformed estimate of heterogeneity in score reliability; Note
that, since in a meta-analysis, we attempt to estimate mean and variance
of the parameter's true distribution, in this table
$\hat{\mu}_{\delta_0}$ represents the meta-analytic estimate of mean
uncorrected $d$ and $\hat{\tau}_{\delta_0}$ represents the estimate of
its heterogeneity; accordingly, $\hat{\mu}_{\delta}$ represents the
estimate of the meta-analytic mean of corrected $d_c$ and
$\hat{\tau}_{\delta}$ the estimate of its heterogeneity.

Secondly, in @tbl-tauES we find that for ten out of twelve phenomena, ES
heterogeneity grows as a result from the attenuation correction
procedure. Even though these differences tend to be small, it is clear
that in all those cases, not only did accounting for score reliability
not explain any ES heterogeneity, we find even more than initially
expected. This is incompatible with claims made in W&D and H&S, the
attenuation correction increased ES heterogeneity. Concerning the
remaining two phenomena, for the Shnabel_Willingness_Reconcile_Rev, the
reduction in heterogeneity is miniscule and the estimate of
heterogeneity is not statistically significant in the first place. Most
likely, the difference in ES heterogeneity estimates can be explained by
issues of estimation, as only eight replications are available.
Similarly, only for the Srull_Behaviour_Hostility phenomenon, we find a
substantial reduction in heterogeneity due to the attenuation correction
procedure. The estimates of ES heterogeneity are not statistically
significant, however, which hampers our ability to meaningfully
interpret and compare these estimates. For the majority of phenomena it
appears that differences in score reliability do not inflate
heterogeneity in uncorrected effect sizes, but instead mask
heterogeneity, that can be unveiled by accounting for score reliability.

@@

In the previous section we demonstrated that assumptions formulated in
the literature regarding meta-scientific methodology are false. We found
a large number of phenomena where, contrary to popular claims, score
reliability does not explain ES heterogeneity. Over the following
paragraphs, we discuss how the true relationship between these two
parameters might look like. Therein, we also identify under which
circumstances an attenuation correction can actually reduce ES
heterogeneity - and where it is more likely to inflate it.

### Hunter & Schmidt's misinterpreted equation

Hunter & Schmidt claim to have derived an equation which demonstrates
how heterogeneity in score reliability inflates heterogeneity in ESs.
“If the level of reliability is independent of the true effect size
across studies, then, to a close approximation:” (p. 309). However, as
can be easily verified, the equation presented in Hunter & Schmidt
(ref.), Equation X on p. 309, does not imply that assumption across all
conditions.

@@

We adjusted @eq-HS using the notation used throughout this text. Here,
$\delta_0$, carrying a 0 in the index, refers to the uncorrected ES,
undistorted by sampling error but not corrected for measuring error. On
the other hand $\delta$, no index, refers to the corrected ES,
undistorted by sampling error and measurement error. Lastly,
$\rho_{XX'}$ refers to score reliability.

@eq-HS implies that heterogeneity in $\delta_0$ is essentially a
function in which heterogeneity and mean value of $\delta$ are weighted
by mean score reliability $\rho_{xx‘}$ and its heterogeneity. Using
selected values for the parameters in @eq-HS, we demonstrate how the
heterogeneity in standardized ES can be both inflated and deflated by
the presence of differences in score reliability. In @fig-hs, the
results of employing @eq-HS to compute the heterogeneity (variance) in
uncorrected ES $\delta_0$ is highlighted. All else held equal, only the
individual level of mean reliability $\mu_{\rho_{xx‘}}$ is varied from
.5 to .8. Mean corrected ES $\mu_{\delta_0}$ is held constant at 1, its
heterogeneity $\tau_{\delta_0}$ held constant at .2 and the
heterogeneity of score reliability $\tau_{\rho_{xx‘}}$ held constant at
.12.

```{r, echo = FALSE}
#| label: fig-HS
#| fig-cap: Implications by Hunter & Schmidt's @eg-HS \n Note: The red line corresponds to the variance of uncorrected ES $\delta_0$.

library(ggplot2)

mu_delta0 <- 1
var_delta0 <- .2^2
mu_alpha <- seq(from = .5, to = .8, length.out = 100)
var_alpha <- .12^2
mu_delta <- mu_delta0/sqrt(mu_alpha)

# var_delta0 <- mu_alpha^2 * var_delta + mu_delta^2 * var_alpha
var_delta <- (var_delta0 - mu_delta^2 * var_alpha) / (mu_alpha^2)

ggplot() +
  geom_hline(yintercept = .04, colour = "red", linewidth = 1, linetype = "dashed") +
  geom_line(aes(y = var_delta, x = mu_alpha), linewidth = 2) +
  theme(panel.background = element_rect(fill = "transparent"), 
        plot.background = element_rect(fill = "transparent", colour = "transparent"), 
        panel.grid.major.y = element_line(colour = "grey"),
        panel.grid.major.x = element_line(colour = "grey"),
        panel.grid.minor = element_line(colour = "transparent"),
        axis.ticks = element_line(colour = "grey"),
        strip.background = element_rect(fill = "transparent")) +
  labs(y = expression(tau^2~"["~delta~"]"),
       x = expression(mu~"["~rho[xx]~"]"))
```

If everything else is held constant, the variance in corrected ES
$\delta$ can be a function only of the mean level of score reliability
$\mu_{\rho_{xx‘}}$, according to H&S equation. In that case, Figure 2
demonstrates that correcting for imperfect score reliability does not
inevitably lead to lower heterogeneity, as the line representing
heterogeneity in corrected ES does cross the red line, representing the
heterogeneity in uncorrected ES. Depending on the mean level of score
reliability, heterogeneity in uncorrected ES can be higher or lower than
heterogeneity in corrected ES.

However, @eq-hs is additionally flawed, as it violates a crucial
definition. As Hunter & Schmidt point out, @eq-hs is only valid “If the
level of reliability is independent of the true effect size across
studies \[…\]” (2014, p. 309). However, if we assume that true score
variance $\sigma^2_T$ varies across administrations, it is unlikely that
this assumption can hold. As demonstrated further above in @eg-dct, ES
$\delta$ is standardized using the true standard deviation $\sigma_T$ in
the sense of CTT. The fact that this parameter is found both in @eq-dct
and @eq-rel demonstrates that ES $\delta$ and score reliability
$\rho_{xx‘}$ are unlikely to constitute truly independent variables, as
long as true score variance is not identical across administrations.

@@

#### Methods

>>>>

Large-scale collaborative replication efforts have sparked discussions
surrounding the replicability of psychological phenomena. For attempts
to estimate a phenomenon’s replicability or to predict whether a future
replication will replicate successfully, heterogeneity is a crucial
parameter that needs to be assessed. In the meta-analytic context,
heterogeneity describes the variation of effect sizes, free of sampling
error. Therefore, the presence of heterogeneity can imply that some
replications of a single phenomenon may be successful, while others are
not. If heterogeneity increases, meaning the phenomenon’s effect size
varies more strongly for unexplained reasons, the probability of
observing an effect size around zero or even in the negative space grows
larger as well. If we know the mean size and heterogeneity of a
phenomenon’s effect, we would theoretically be able to establish an
expected replication rate (reference).

Similarly, it has been argued that heterogeneity in effect sizes is an
indicator of the theory’s “completeness” surrounding the phenomenon.
Linden and Hönekopp argue that “low (as opposed to high) heterogeneity
reflects a more advanced understanding of the subject matter being
studied” (2021, p.2). Similarly, Schuetze and von Hippel (2023) argue
that heterogeneity in effects is an indicator of a vague, poorly
specified theory.

Large-scale attempts of direct replications, using identical protocols,
such as the Many Labs studies or Registered Replication Reports
(references), for the first time allow researchers to estimate
heterogeneity undistorted by selection processes like publication bias.
In re-analyses of these studies, Olsson-Collentine et al. (2020)
identify a positive correlation between a phenomenon’s effect size and
its heterogeneity. Similarly, van Erp et al. (2017) and Stanley et al.
(2018) estimate strong degrees of heterogeneity across conceptual
replications in Psychology. In a separate re-analysis of large scale
direct replications, Renkewitz et al. (in preparation) identify
substantial heterogeneity in almost all projects where a non-zero effect
could be identified. This aligns with the correlation found by
Olsson-Collentine et al. (2020).

One source for this heterogeneity, which is not a typical moderator tied
to a specific theory, might be impairments in score reliability which
vary in their extent across replications. This is an argument put
forward repeatedly by original researchers and replicators alike (refs).
Hunter and Schmidt (xxx) and Wiernik and Dahlke (xxx), discussing the
role of score reliability for standardized effect sizes, claim that
differences in score reliability should inflate, and therefore increase,
heterogeneity. This implies that standard corrections for imperfect
score reliability, such as attenuation correction procedures, should
remove that inflation, reducing heterogeneity in corrected effect sizes.
Over the following pages, we assess whether score reliability can in
fact be made responsible for heterogeneity identified across phenomena.
We find no empirical evidence for the argument that correcting for
differences in score reliability reduces effect size heterogeneity.
Instead, to our great surprise, we find that heterogeneity might as well
be larger after the correction procedure. Therefore, theoretical
analyses regarding the relationship between score reliability and
heterogeneity need to be false. In the second half of this text we make
use of analytical arguments to identify the necessary conditions under
which an attenuation correction inevitably leads to an increase in
effect size heterogeneity.

@@

Here, MD refers to the mean difference between two groups of interest,
while $\sigma_X$ is the total pooled standard deviation. Since $d$
constitutes an observed effect size it is affected by sampling error. In
terms of heterogeneity, we are not interested in the variance of $d$,
but the variance of the true, underlying ES $\delta$. In a
random-effects meta-analytic model, we can use the standard error and an
estimate of heterogeneity to understand how far off the individual
observed $d$ is from the true $\delta$ in a weighting procedure
(reference). The standard error of Cohen’s $d$ is computed as defined in
@eq-SEd.

@@

Therefore, we assess the extent to which this intuition is supported by
empirical evidence in the following section. The re-analysis of archival
data serves to shed light on the following research question:

-   To what extent can heterogeneity in standardized effect sizes be
    explained by means of an attenuation correction?

@@

The ManyLabs projects were collaborative efforts to replicate several
psychological phenomena across different research sites, employing
identical protocols. From five published projects, the data for the
first three and the fifth Many Labs projects was harmonized and publicly
stored in a standardized format when the DRIPHT repository was set up
(references). The Registered Replication Reports are similar
collaborative efforts, with the sole distinction that for each report a
single phenomenon was replicated several times. The Registered Reports
3-10 were added to the DRIPHT repository, as they employed experimental
designs with at least two groups. Lastly, some projects from the
Psychological Science Accelerator were added to the DRIPHT repository.
In contrast to Many Labs or Registered Replication Projects, these
collaborative efforts, while also distributed across the globe, do not
formally attempt to perform direct replications, but focus on original
research or conceptual replications. However, as each project used the
same protocol and participants are distributed across different
countries, a data structure similar to that of a multi-site replication
study emerges. What makes all these collaborative efforts valuable for
this manuscript is the fact that all phenomena collected in the
DRIPHT-repository are making use of two-group designs and are therefore
easily assessed using standardised mean differences.

While replication data on a decent number of psychological phenomena is
available in the DRIPHT-repository, to demonstrate how incorporating
score reliability affects heterogeneity in ES, selected phenomena need
to fulfil three conditions: (1) As we focus on the use of ES $d$, the
phenomenon needs to be assessed using a group design consisting of at
least two groups. (2) Score reliability estimates can be derived, this
implies that the phenomenon needs to be measured using either several
indicators forming a single scale or by repeatedly measuring across
several timepoints. (3) Score reliability can only attenuate effect
sizes that are not zero in the first place. Therefore, we focus on
phenomena where meta-analytic mean ES can be statistically distinguished
from zero. While the first criterion is fulfilled by all phenomena found
in the DRIPHT repository, the second and third criteria still need
evaluation.

As all phenomena in the collection fulfil the first condition, 50
phenomena have been catalogued where the effect is studied by comparing
the outcome across two separate groups. From the collection of 50
phenomena, 21 fulfil the second condition, so that estimates of score
reliability can be derived. For the 21 phenomena, the dependent variable
used to construct the effect size is measured by employing more than a
single indicator. The number of indicators per scale varied between 3
and 30. The majority of scales was made up of 3 to 6 items. At the same
time, however, this implies that the remaining 29 phenomena did not make
use of more than a single indicator to measure the dependent variable.
Therefore, for the majority of phenomena it is not possible to assess in
how far measurement quality is sufficiently high in terms of e.g.
internal consistency or low random error variance.

From the 21 phenomena that employed more than a single indicator, 19
made use of Likert-style items, where respondents would indicate their
agreement to some kind of statement. The scale-length varied from 3 to
10-points. The items measuring the remaining 2 phenomena were coded
dichotomously, as responses were right or wrong. While the majority of
designs make use of some form of control-treatment manipulation,
randomly assigning participants to a condition (19 phenomena), some
phenomena refer to the effect of pre-existing differences, e.g.
biological sex (2 phenomena). Subsequently, those phenomena need to be
identified where the meta-analytic mean effect size is statistically
significant different from zero. This will be done as a preliminary step
in this manuscript’s analysis procedure. Detailed information concerning
the different phenomena and how they were measured can be found at
\[osf-link\], while a brief summary can be found in Table 1.

@@

For each replication of a phenomenon, Cohen’s $d$ is computed as an
estimate of raw ES, according to @eq-d, including its standard error
(@eq-SEd). Subsequently, a random-effects meta-analysis is performed
using metafor version X.XX. To estimate meta-analytic mean ES and
heterogeneity we make ues of the REML-estimator (ref). To identify which
phenomena pass criterion (3), using a Wald-type significance test we
assess for which phenomenon the meta-analytic mean ES is significantly
different from zero. For this hypothesis test, as all other hypothesis
tests in this manuscript, we employ a one-tailed significance level of
.05.

Additionally, estimates of score reliability are derived. While not
without criticism (references), Cronbach’s Alpha is used to estimate
score reliability. It is often noted that tau-equivalence is an
unrealistic assumption to hold against real-world data, implying that
other estimates of score reliability, such as McDonald’s Omega
(reference), Guttman’s Lambda 2 or 4 (reference), or the Greatest Lower
Bound (reference) might be better suited. However, all scales employed
in this project are scored by computing the simple mean or sum of
responses across items for each individual. Computing the mean or sum of
several items implicitly assumes tau-equivalence, as each item
contributes equally to the individual’s test score. Therefore,
Cronbach’s Alpha is the better choice to estimate score reliability, as
it avoids a mismatch in assumptions between how the score is computed
and how score reliability is estimated. The individual estimates of
Cohen’s d are corrected using these estimates of Cronbach's Alpha,
according to @eq-dc, including the corrected standard errors (@eq-SEdc).
Thereby, estimates of ES are computed, which are corrected for imperfect
reliability. Again, a random-effects meta-analysis is run using metafor,
generating an estimate of heterogeneity in corrected ES. Different
indicators of heterogeneity are readily available, such as $I^2$, $H^2$
or the coefficient of variation. However, the descriptions found in H&S
and W&D discuss the absolute amount of heterogeneity in terms of
variance ($\tau^2$) or standard deviation ($\tau$). These parameters
discuss the variability of the standardized ES in the population in
terms of variance or standard deviation. Therefore, we also make use of
absolute heterogeneity here.

Subsequently, in order to assess whether ES heterogeneity was indeed
reduced by the attenuation correction procedure, the estimates of
heterogeneity in uncorrected ES and corrected ES are compared. Using
Cochran’s *Q*-test, we assess whether the estimates of heterogeneity are
statistically significantly different from 0. Only for those projects
where we have sufficient confidence that the individual estimates of
heterogeneity are larger than zero can we actually begin to interpret
whether there was any change.

@@

We executed these steps for each phenomenon. The statistical programming
language R, Version X.XXX (reference) is used for all data manipulation
and statistical analysis.

### Results

[Equations @eq-d], [-@eq-SEd], [-@eq-dc], and [-@eq-SEdc] were used to
generate estimates of standardized ES, both uncorrected and corrected,
with their corresponding standard errors. Generally, this leads to
larger (absolute) ES and larger standard errors after the correction. As
an example from the 21 selected phenomena, the results from this
procedure on Nosek’s Explicit Art sex differences phenomenon are
displayed in the forest plot in @fig-forest. Here, grey dots represent
the uncorrected ES for each sample, while black dots represent the
corrected ES in each sample. The bars surrounding these dots show the
respective 95%-Cis.

```{r, echo = FALSE, fig.height = 6}
#| label: fig-forest 
#| #| fig-cap: Forest Plot Nosek_Explicit_Art 
forest_plot_rel(ES_rma_df[which(ES_rma_df$MASC == "Nosek_Explicit_Art"),], agg_L[[which(MASC_names == "Nosek_Explicit_Art")]])
```

@@

#### Which phenomena pass criterion 3)?

@tbl-meanES demonstrates that what we observed in @fig-forest holds
across all 21 phenomena. All meta-analytic effects are larger in their
absolute size after applying an attenuation correction procedure.
Similarly, the uncertainty quantified in the estimate's standard error
is larger. Additionally, @tbl-meanES highlights which phenomena pass
criterion (3)—the effect size must be statistically significantly
distinguishable from zero. 12 phenomena pass this criterion, highlighted
in boldface. It may be remarked that the attenuation correction
procedure, in this data, appears to have no influence on whether a
phenomenon passes the criterion. The differences in p-value are rather
small and lead to no difference in conclusion, regarding a significance
level

```{r, echo = FALSE}
#| label: tbl-meanES
#| tbl-cap: Meta-analytic average  ES
knitr::kable(x = read.csv(here("Tables/Mean_ES.csv")),
             col.names = c("MASC", 
                           "$\\hat{\\mu}_{\\delta_0}$ (SE)", 
                           "p", 
                           "$\\hat{\\mu}_{\\delta}$ (SE)", 
                           "p"),
             escape = FALSE)
```

*Note:* Since in a meta-analysis, we attempt to estimate mean and
variance of the parameter's true distribution, in this table
$\hat{\mu}_{\delta_0}$ represents the meta-analytic estimate of mean
uncorrected $d$ and $\hat{\mu}_{\delta}$ represents the estimate of the
meta-analytic mean of corrected $d_c$.

#### Estimates of heterogeneity

@tbl-tauES contains the results from performing an RG-MA on the
Cronbach's Alpha coefficients of the scales used to measure the 12
remaining phenomena. Additionally, it contains estimates of
heterogeneity for both uncorrected ES ($\delta_0$) and corrected ES
($\delta$), which are central to assess the research questions posed in
section XXX. Most importantly, @tbl-tauES demonstrates that for all but
one of the phenomena, the attenuation correction procedure barely has an
impact on the extent of ES heterogeneity. However, while the
heterogeneity for the Srull_Behaviour_Hostility phenomenon seems reduced
by about a third of its initial value, the ES heterogeneity for this
phenomenon is rather small to begin with. Using a significance test, we
can not distinguish this value confidently from zero. This implies that
for the majority of phenomena, differences in score reliability can not
be made responsible for heterogeneity in standardized effect sizes.

@@

*Note:* $\hat{\mu}_{\rho_{XX'}}$ is the back-transformed meta-analytic
estimate of mean score reliability; $\tau{\mu}_{\rho_{XX'}}$ is the
back-transformed estimate of heterogeneity in score reliability; Note
that, since in a meta-analysis, we attempt to estimate mean and variance
of the parameter's true distribution, in this table
$\hat{\tau}_{\delta_0}$ represents the estimate of heterogeneity in
uncorrected $d$; accordingly, $\hat{\tau}_{\delta}$ represents the
estimate of the heterogeneity in corrected $d_c$.

Secondly, in @tbl-tauES we find that for ten out of twelve phenomena, ES
heterogeneity grows as a result from the attenuation correction
procedure. Even though these differences tend to be small, it is clear
that in all those cases, not only did accounting for score reliability
not explain any ES heterogeneity, we find even more than initially
expected. This is incompatible with claims made in W&D and H&S, the
attenuation correction increased ES heterogeneity. Concerning the
remaining two phenomena, for the Shnabel_Willingness_Reconcile_RPP, the
reduction in heterogeneity is miniscule and the estimate of
heterogeneity is not statistically significant in the first place. Most
likely, the difference in ES heterogeneity estimates can be explained by
issues of estimation, as only eight replications are available.
Similarly, only for the Srull_Behaviour_Hostility phenomenon, we find a
substantial reduction in heterogeneity due to the attenuation correction
procedure. The estimates of ES heterogeneity are not statistically
significant, however, which hampers our ability to meaningfully
interpret and compare these estimates. For the majority of phenomena it
appears that differences in score reliability do not inflate
heterogeneity in uncorrected effect sizes. Instead these differences
mask ES heterogeneity, that can be unveiled by accounting for score
reliability.

@@

In the re-analysis of archival data in the previous section we
demonstrated that, at least for the multi-site direct replication and
PSA-projects, that score reliability does not hold any noteworthy value
as a potential moderator explaining ES heterogeneity. Additionally, we
demonstrated that the current consensus formulated in the literature
regarding how score reliability increases ES heterogeneity is false. We
found a large number of phenomena where, contrary to popular claims,
correcting for score reliability increases ES heterogeneity, instead of
decreasing it.

This implies that there is an error in the reasoning or intuition behind
claims made in H&S and W&D, as our observations can not be explained by
the current literature. Over the following paragraphs, we will briefly
explore the intuition that might have led to the previous claims.
Subsequently, we will use analytical arguments to clarify how score
reliability actually affects ES heterogeneity. Therein, we also identify
potential conditions under which an attenuation correction can actually
reduce ES heterogeneity - and under which it is more likely to increase
it.

### Intuitive relationship

It might help to think about linear models, when discussing the
intuition that might underlie the idea, that controlling for score
reliability should reduce ES heterogeneity. Heterogeneity describes
unexplained variation in a parameter's true distribution, that we model
using meta-analytical techniques. In that way, the idea of heterogeneity
is closely related to residual variance, as we observe it in a simple
linear model. Both describe variation that could not be explained by the
predictive variables in the case of the simple linear model, or not
explained by incorporating the Standard Errors in the case of a
meta-analytic model.

For the simple linear model, if we add an additional predictor to the
model that we know shares variance with the parameter that we aim to
predict, this should reduce the amount of residual variance that we
observe across most instances. Generally, removing a source of variance
by controlling for it, we expect any unexplained variation to shrink.
Similarly, we know that differences in score reliability, by definition,
are related to differences in standardized effect sizes. Therefore,
intuitively, it seems sensible to expect that controlling for
differences in score reliability should reduce the unexplained variance
in standardized effect sizes, as described in H&S and W&D.

Hunter & Schmidt have additionally derived an equation which
demonstrates how heterogeneity in score reliability affects
heterogeneity in ESs.\@eq-HS implies that heterogeneity in $\delta_0$ is
essentially a function in which heterogeneity and mean value of $\delta$
are weighted by mean score reliability $\rho_{xx‘}$ and its
heterogeneity.

@@

We adjusted @eq-HS using the notation used throughout this text. Here,
$\delta_0$, carrying a 0 in the index, refers to the uncorrected ES,
undistorted by sampling error but not corrected for measuring error. On
the other hand $\delta$, no index, refers to the corrected ES,
undistorted by sampling error and measurement error. Lastly,
$\rho_{XX'}$ refers to score reliability. Hunter & Schmidt (xxx) claim
that @eq-HS demonstrates that differences in score reliability inflate
ES heterogeneity, which means that an attenuation correction procedure
would have to reduce ES heterogeneity. However, it can be easily
verified that @eq-HS does not allow for such claims. Depending on both
the mean value $\mu_{\rho_{xx‘}}$ and the heterogeneity
$\tau^2_{\rho_{xx‘}}$ in score reliability, ES heterogeneity might be
increased or decreased by the differences in score reliability.

Most likely, Hunter and Schmidt did not fully explore the implications
of @eq-HS, but followed the general intuition we lined out–"If a source
of variation is controlled for, the amount of unexplained variance
should be reduced." Wiernik and Dahlke (ref) refer to the work by Hunter
& Schmidt (ref), likely following their intuition.

In the re-analysis of archival data, we have demonstrated that this
intution does not hold for the relationship of ES heterogeneity and
score reliability. Instead, we will propose an alternative way that
helps us discern, under which conditions we can expect an attenuation
correction procedure to lead to an increase in ES heterogeneity.

@@

#### Methods {#sec-methods2}
